### 2016-05-30 thesis/methods/methods.tex
tmp commit

%TODO: change the title name?
\section{\gls{icgc} sample exclusion criteria}
\label{sec:sampleexclusion}

@@

\subsection{Sample BMI availability}
\label{subsec:samplebmiavailability}

@@

\subsection{Sample ID conversion and matching}
\label{subsec:sampleidconversion}

@@

@@

@@

Each of the eight cancer data from \cref{subsec:datprocicgc} were transformed using the transformation matrix from \cref{subsec:transmatcr}.
Metagene scores of the samples in each cancer types were obtained by taking the first column of the cancer matrix that resulted from the transformation.
The \gls{metagene} scores were ranked based on the number of samples in the data, and the scores ranged from 0 to 1.

\subsection{Creighton metagene and sample gene expression in \gls{icgc} data}
\label{subsec:crmetaicgcge}

\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap for each cancer type.
The data from \cref{subsec:datprocicgc} were used for the heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Creighton metagene and sample \gls{bmi} in \gls{icgc} data}
\label{subsec:crmetaicgcge}

@@

\subsection{Fuentes-Mattei metagene and gene expression pattern in Creighton data}
\label{subsec:fmmetage}

@@

@@

\subsection{Metagene and sample \gls{bmi}}
\label{subsec:fmmetabmi}

@@

\subsection{Fuentes-Mattei metagene and sample gene expression in \gls{icgc} data}
\label{subsec:fmmetaicgcge}

@@

\subsection{Creighton metagene and sample \gls{bmi} in \gls{icgc} data}
\label{subsec:fmmetaicgcge}

@@

The difference in the gene expression between the obese and non-obese sample groups in each of the four data sets described in \cref{subsec:subdatadeg} were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
From the top significantly expressed genes produced by the \texttt{topTable} function, all of the gene probes that had a p-value \textless{} 0.01 were selected.

When there were more than 799 gene probes in the list of significant genes, the top 799 gene probes were selected from the list and used as the gene probe set for the metagene analysis, for that dataset.

For each gene probe set, the gene probes that were common with the obesity associated genes identified by \citet{Creighton2012} were selected and used as a separate gene probe set.
There were eight metagenes in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from \citet{Creighton2012} study.

\subsection{Gene probe set conversion}
\label{subsec:gpconv}

\subsection{BMI metagenes in Creighton data}
\label{subsec:bmidegmetacr}

@@

The gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.

>>>>

%TODO: change the title name?
\subsection{\gls{icgc} sample exclusion criteria}
\label{subsec:sampleexclusion}

@@

\subsubsection{\textbf{Sample BMI availability}}
\label{subsubsec:samplebmiavailability}

@@

\subsubsection{\textbf{Sample ID conversion and matching}}
\label{subsubsec:sampleidconversion}

@@

\subsection{Creighton metagene generation}
\label{subsec:crmetagen}

@@

\Gls{metagene} scores from \cref{subsec:crmetagen} were plotted together with the gene expression of the obesity associated gene probes for each sample on a heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

@@

Each of the eight cancer data from \cref{subsec:datprocicgc} were transformed using the transformation matrix from \cref{subsec:transmatcr}.
Metagene scores of the samples in each cancer types were obtained by taking the first eigenvector of the cancer matrix that resulted from the transformation.
The \gls{metagene} scores were ranked based on the number of samples in the data, and the scores ranged from 0 to 1.

\subsection{Creighton metagene and sample gene expression}
\label{subsec:crmetaicgcge}

\Gls{metagene} scores from \cref{subsec:icgctranscr} were plotted together with the gene expression of the obesity associated gene probes on a heatmap for each cancer type.
The data from \cref{subsec:datprocicgc} were used for the heatmap.
The samples in the heatmap were ordered based on their ranked \gls{metagene} scores.
To have a better colour scales in the heatmap, any expression values that were greater than 3 or less than -3 were assigned the value of 3 and -3, respectively.

\subsection{Creighton metagene and sample \gls{bmi}}
\label{subsec:crmetaicgcge}

@@

\subsection{Fuentes-Mattei metagene generation}
\label{subsec:fmmetagen}

@@

\subsection{Fuentes-Mattei metagene and gene expression in Creighton data}
\label{subsec:fmmetage}

@@

\subsection{Fuentes-Mattei metagene and sample \gls{bmi} in Creighton data}
\label{subsec:fmmetabmi}

@@

\subsection{Fuentes-Mattei metagene and sample gene expression}
\label{subsec:fmmetaicgcge}

@@

\subsection{Creighton metagene and sample \gls{bmi}}
\label{subsec:fmmetaicgcge}

@@

The difference in the gene expression between the obese and non-obese sample groups in each of the four data sets described in \cref{subsec:subdatadeg} were analysed by fitting a linear model to the data, using the \textit{limma} package in \gls{r}.
From the list of significantly expressed (genes produced by the \texttt{topTable} function), all of the gene probes that had a p-value \textless{} 0.01 were selected.

When there was more than 799 gene probes in the list of significant genes, the top 799 gene probes were selected from the list.
This list of 799 gene probes represented the obesity associated metagene for that dataset.

For each gene probe set, the gene probes that were common with the obesity associated genes identified by \citet{Creighton2012} were selected and used as a separate gene probe set.
There were eight metagenes in total after this process: four gene probe sets from the datasets from \cref{subsec:subdatadeg}, and four gene probe sets that were in common with the obesity associated genes from the \citet{Creighton2012} study.

\subsection{Gene probe set conversion and selection}
\label{subsec:gpconv}

For each of the eight gene probe sets identified in \cref{subsec:crdeg}, the gene probe IDs were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Any duplicates and any gene symbols that were not present in both the Creighton data and the ICGC data were removed from the gene probe set.

\section{BMI metagenes in Creighton data}
\label{sec:bmidegmetacr}

\subsection{BMI metagene generation}
\label{subsec:bmimetagen}

\subsection{BMI metagenes and sample gene expression in Creighton data}
\label{subsec:gpconv}

The gene probe IDs in the Creighton data were converted into gene symbols using the \textit{hgu133a.db} package in \gls{r}.
Where there was a duplicated gene in the data, appropriate gene data was chosen to represent that particular gene by using the \texttt{collapseRows} function in the \textit{WGCNA} package in \gls{r}.

For each BMI gene probe set identified in \cref{sec:deg}, relevant genes were taken from the Creighton data.
The data was then standardised so all of the genes had a mean ($\mu$) of 0 and standard deviation ($\sigma$) of 1.

@@
